Make any stupid plot you want

### 2021-05-05 draft/sections/math.tex
graphics section

\subsubsection{Continuity of the Graphic \gbase} 
\label{sec:math:graphic:base}
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/retraction_maps.png}
    \caption{For a visualization component to preserve continuity, it must have a continuous surjective map $\vindex: \gbase\rightarrow\dbase$ from graphic continuity to data continuity. The scatter (a) and line (b) graphic base spaces \gbase\ have one more dimension of continuity than \dbase\ so that \gbase\ can encode physical aspects of the glyph, such as shape (a circle) or thickness. The image (c) has the same dimension in \gbase\ as in \dbase\ because \dbase\ is already 2D and therefore can directly map into screen space.}
    \label{fig:math:graphic:retraction:map}
\end{figure}
To establish that a visualization component preserves continuity, we propose that their must be a continuous map $\vindex: \gbase\rightarrow\dbase$ from the graphic base space to the data space. For example, consider a \gbase\ that is mapped to the region of a 2D display space that represents \dbase. For some visualizations, \dbase\ may be lower dimension than \gbase. For example, a point that is 0D in \dbase\ cannot be represented on screen unless it is thickened to 2D to encode the connectivity of the pixels that visually represent the point. This thickening is often not necessary when the dimensionality of \dbase\ matches the dimensionality of the target space, for example if \dbase\ is 2D and the display is a 2D screen. We introduce \gbase\ to thicken \dbase\ in a way which preserves the structure of \dbase.

Formally, we require that \dbase\ be a deformation retract\cite{RetractionTopology2020} of \gbase\ so that \dbase\ and \gbase\ have the same homotopy, meaning there is a continuous map from \gbase\ to \dbase\cite{weissteinHomotopy}. The surjective map $\vindex: \gbase \rightarrow \dbase$ 
\begin{equation}
    \begin{tikzcd}
        \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
        \dbase                   & \gbase \arrow[l, "\vindex"']
    \end{tikzcd}
    \label{eq:math:graphic:vindex}
\end{equation}
goes from region $\gbasepoint \in \gbase_{\dbasepoint}$ to its associated point $\gbasepoint$. This means that if $\vindex(\gbasepoint) = \dbasepoint$, the record at \dbasepoint\ is copied over the region \gbasepoint\ such that $\dsection(\dbasepoint)=\vindex^*\dsection(\gbasepoint)$ where $\vindex^*\dsection(\gbasepoint)$  is \dsection\ pulled back over \gbase. The map \vindex\ is part of the implementation of the artist $\mathcal{\vartist}$ and therefore is not defined in terms of the data; instead it is how we specify the constraint that the type of the graphic \textit{continuity} must be able to map to the type of the data \textit{continuity}. When \dbase\ is discrete points and the graphic is a scatter plot, each point $\dbasepoint \in \dbase$ can correspond to a 2D disk $\gbase_{\dbasepoint}$ as shown in \autoref{fig:math:graphic:retraction:map}. In the case of 1D continuous data and a line plot, the region \gy\ over a point $\gx_i$ specifies the thickness of the line in \gbase\ for the corresponding \dsection\ on \dbasepoint. The image has the same dimensions in data space and graphic space such that no extra dimensions are needed in \gbase.

The mapping function \vindex\ provides a way to identify the part of the visual transformation that is specific to the the connectivity of the data rather than the values; for example it is common to flip a matrix when displaying an image. The \vindex\ mapping is also used by interactive visualization components to look up the data associated with a region on screen.  One example is to fill in details in a hover tooltip, another is to convert region selection (such as zooming) on \gbase\ to a query on the data to access the corresponding record components on \dbase.

\subsubsection{Graphic \gsection}
\label{sec:graphic_section}
The section $\gsection: \gbase \rightarrow \gtotal$ is the graphic in an idealizes prerender space and also acts as a specification for rendering the graphic to an image. It is sufficient to sketch out how an arbitrary pixel would be rendered, where a pixel $p$ in a real display corresponds to a region $\gbase_p$ in the idealized display. To determine the color of the pixel, we aggregate the color values over the region via integration:
\begin{align*}
    \label{eq:math:graphic:section:color}
    r_p &= \iint\limits_{S_p} \rho_r(s)ds^{2}\\
    g_p &= \iint\limits_{S_p} \rho_g(s)ds^{2}\\
    b_p &= \iint\limits_{S_p} \rho_b(s)ds^{2}
\end{align*}
For a 2D screen, the pixel is defined as a region $p=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]$ of the rendered graphic. Since the x and y in $p$ are in the same coordinate system as the x and y components of \gfiber\,  the inverse map of the bounding box $\gbase_{p} ={\gsection_{xy}}^{-1}(p)$ is a region $\gbase_p \subset \gbase$. The color is the result of the integration over  $\gbase_p$.

\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/render.png}
    \caption{To render a graphic, a pixel $p$ is selected in the display space, which is defined in the same coordinates as the x and y components in \gfiber\ via the renderer. In \gtotal\, the inverse mapping ${\gsection_{xy}}^(p)$ returns a region $\gbase_{p} \subset \gbase$. $\gsection(\gbase_{p})$ returns a set of points $(x,\,y,\,r,\,g,\,b) \in \gfiber$ that lie over $\gbase_{p}$. The integral over the $(r,\,g,\,b)$ pixels specifies that the pixel should be green}
    \label{fig:math:graphic:rho:lookup}
\end{figure}

As shown in \autoref{fig:math:graphic:rho:lookup}, a pixel $p$ in the output space, drawn in yellow, is selected and mapped, via the renderer, into a region on \gtotal.  The region on \gtotal\ corresponds to a region $\gbase_p \subset \gbase$ via the inverse mapping  ${\gsection_{xy}}^(p)$. The base space \gbase\ is an annulus to match the topology of the graphic idealized in \gtotal. The section  $\gsection(\gbase_{p})$  then maps into the fiber \gfiber\ over $\gbase_p$ to obtain the set of points in \gfiber, here represented as a table, that correspond to that section. The integral over the pixel components of this set of points in the fiber yields \{3, 135, 60\} the actual color of the pixel. In general, \gsection\ is an abstraction of rendering. In very broad strokes \gsection\  can be a specification such as PDF\cite{bienz1993portable}, SVG\cite{quintScalable2003}, or an openGL scene graph\cite{CarsonOpenGL1997}. Alternatively, \gsection\ can be a rendering engine such as cairo\cite{CairographicsOrg} or AGG\cite{shemanarevAntiGrainGeometry}. Implementation of \gsection\ is out of scope for this work,

>>>>

\subsubsection{Continuity of the Graphic \gbase} 
\label{sec:math:graphic:base}

For a visualization component to preserve continuity, we propose that there must exist a structure preserving surjective map $\vindex: \gbase\rightarrow\dbase$ from the data base space \dbase\ to the graphic base space \gbase. Formally, we require that \dbase\ be a deformation retract\cite{RetractionTopology2020} of \gbase\ such that \dbase\ and \gbase\ have the same homotopy, meaning there is a continuous map from \gbase\ to \dbase\cite{weissteinHomotopy}. The surjective map $\vindex: \gbase \rightarrow \dbase$ 
\begin{equation}
    \begin{tikzcd}
        \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
        \dbase                   & \gbase \arrow[l, "\vindex"']
    \end{tikzcd}
    \label{eq:math:graphic:vindex}
\end{equation}
goes from region $\gbasepoint \in \gbase_{\dbasepoint}$ to its associated point $\dbasepoint \in \dbase$. This means that if $\vindex(\gbasepoint) = \dbasepoint$, the record at \dbasepoint\ is copied over the region \gbasepoint\ such that $\dsection(\dbasepoint)=\vindex^*\dsection(\gbasepoint)$ where $\vindex^*\dsection(\gbasepoint)$  is \dsection\ pulled back over \gbase. The map \vindex\ is part of the implementation of the artist $\mathcal{\vartist}$ and therefore is not defined in terms of the data; instead it is how we specify the constraint that the type of the graphic \textit{continuity} must be able to map to the type of the data \textit{continuity}.

\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/retraction_maps.png}
    \caption{For a visualization component to preserve continuity, it must have a continuous surjective map $\vindex: \gbase\rightarrow\dbase$ from graphic continuity to data continuity. The scatter (a) and line (b) graphic base spaces \gbase\ have one more dimension of continuity than \dbase\ so that \gbase\ can encode physical aspects of the glyph, such as shape (a circle) or thickness. The image (c) has the same dimension in \gbase\ as in \dbase\ because \dbase\ is already 2D and therefore can directly map into screen space.}
    \label{fig:math:graphic:retraction:map}
\end{figure}

To encode the continuity of the elements in the display fiber \gfiber, the graphic base space \gbase\ has the same dimensionality as the target output space. For example, in \autoref{fig:math:graphic:retraction:map} the base space \gbase\ is a representation of a region of a 2D display space. Since \gbase\ must have the same dimensionality as the output graphic, it is allowed to add dimensions to \dbase\ to make \dbase\ renderable. A point that is 0D in \dbase\ cannot be represented on screen unless it is thickened to 2D (a) to encode the connectivity of the pixels that visually represent the point. This is also the case with the line (b), which would be infinitely think on screen if \gbase\ was not thickened to 2D. This thickening is often not necessary when the dimensionality of \dbase\ matches the dimensionality of the target space, for example if \dbase\ is 2D and the display is a 2D screen (c). Since the mapping function \vindex\ binds the graphic base space to the data base space, it can be used by interactive visualization components to look up the data associated with a region on screen.  One example is to fill in details in a hover tooltip, another is to convert region selection (such as zooming) on \gbase\ to a query on the data to access the corresponding record components on \dbase.

\subsubsection{Graphic \gsection}
\label{sec:math:graphic:section}
The section $\gsection: \gbase \rightarrow \gtotal$ is the graphic in an idealized prerender space and also acts as a specification for rendering the graphic to target display format. To demonstrate the role of \gsection\, it is sufficient to sketch out how an arbitrary pixel would be rendered, where a pixel $p$ in a real display corresponds to a region $\gbase_p$ in the idealized display. To determine the color of the pixel, we aggregate the color values over the region via integration:
\begin{align*}
    \label{eq:math:graphic:section:color}
    r_p &= \iint\limits_{S_p} \rho_r(s)ds^{2}\\
    g_p &= \iint\limits_{S_p} \rho_g(s)ds^{2}\\
    b_p &= \iint\limits_{S_p} \rho_b(s)ds^{2}
\end{align*}
For a 2D screen, the pixel is defined as a region $p=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]$ of the rendered graphic. Since the x and y in $p$ are in the same coordinate system as the x and y components of \gfiber\,  the inverse map of the bounding box $\gbase_{p} ={\gsection_{x,y}}^{-1}(p)$ is a region $\gbase_p \subset \gbase$. The color is the result of the integration over  $\gbase_p$.

\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/render.png}
    \caption{To render a graphic, a pixel $p$ is selected in the display space, which is defined in the same coordinates as the x and y components in \gfiber\ via the renderer. In \gtotal\, the inverse mapping ${\gsection_{x,y}}^(p)$ returns a region $\gbase_{p} \subset \gbase$. $\gsection(\gbase_{p})$ returns a set of points $(x,\,y,\,r,\,g,\,b) \in \gfiber$ that lie over $\gbase_{p}$. The integral over the $(r,\,g,\,b)$ pixels specifies that the pixel should be green}
    \label{fig:math:graphic:rho:lookup}
\end{figure}

As shown in \autoref{fig:math:graphic:rho:lookup}, a pixel $p$ in the output space, drawn in yellow, is selected and mapped, via the renderer, into a region on \gtotal.  The region on \gtotal\ corresponds to a region $\gbase_p \subset \gbase$ via the inverse mapping  ${\gsection_{xy}}^(p)$. The base space \gbase\ is an annulus to match the topology of the graphic idealized in \gtotal. The section  $\gsection(\gbase_{p})$  then maps into the fiber \gfiber\ over $\gbase_p$ to obtain the set of points in \gfiber, here represented as a table, that correspond to that section. The integral over the pixel components of this set of points in the fiber yields the color of the pixel. In general, \gsection\ is an abstraction of rendering. In very broad strokes \gsection\ can be a specification such as PDF\cite{bienz1993portable}, SVG\cite{quintScalable2003}, or an openGL scene graph\cite{CarsonOpenGL1997} or a rendering engine such as cairo\cite{CairographicsOrg} or AGG\cite{shemanarevAntiGrainGeometry}. Implementation of \gsection\ is out of scope for this proposal.
